---
title: "Fungal Guild Assignments"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
author: "Briana K. Whitaker"
date: "`r Sys.Date()`"
---
\fontsize{9}{10}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning =FALSE, message=FALSE)
```
---




* Run using `r version[['version.string']] `.

# Objective
This document reports on the **fungal guild assignments** from the NC Microbiome Landscapes Project, performed in the summer of 2019. 




# 0) Load Packages, set path to data
```{r, echo=FALSE, results='hide', include=FALSE} 

x<-c("ggplot2", "phyloseq", "dplyr", "RColorBrewer", "DESeq2", "devtools",
     "reshape2",  "viridis", "tidyr", "tidyverse")
lapply(x, require, character.only = TRUE)

#Load functions
source("./code/multiplot.function.R")

#add 'not in' function
`%nin%` = Negate(`%in%`)

#set seed
set.seed(829)


# set ggplot2 theme
theme_set(theme_bw(base_size=16)) 
theme_update(panel.grid.major=element_line(0), panel.grid.minor=element_line(0))

# color palettes
display.brewer.pal(8, "Dark2") #assumes Phal is first in order
brewer.pal(8, "Dark2")
spp_color <- c("#E6AB02", "#D95F02", "#1B9E77", '#7570B3') 
#gold, orange, green, purple

#levels(sample_data(ps_filter)$species)
#"CRN"  "SOY"  "SWCH" "WHT" 
#https://mdigi.tools/color-shades/#a6761d
crn_color <- c('#fed561', '#fdc422', '#dda402', '#9e7501')
soy_color <- c('#fea561', '#fd8122', '#dd6102', '#9e4501')
swc_color <- c('#77e8c6', '#40deaf', '#21bf90', '#178867')
wht_color <- c('#cfcde4', '#8f8ac1', '#555096', '#2f2c53')

site_color <- c(crn_color, soy_color, swc_color, wht_color)

```

```{r, include = FALSE}
tidy_psmelt <- function(physeq) {
    # First, make sure taxa are rows
    if (taxa_are_rows(physeq)) {
        st <- otu_table(physeq) %>% t
    } else {
        st <- otu_table(physeq)
    }
    # Convert to a tibble in tidy form
    st <- st %>% 
        as("matrix") %>%
        as_tibble(rownames = "Sample") %>%
        gather("OTU", "Abundance", -Sample)
    # Get the sample data as a tibble
    sam <- sample_data(physeq) %>% 
        as("data.frame") %>% 
        as_tibble(rownames = "Sample")
    # Same for the tax table
    tax <- tax_table(physeq) %>%
        as("matrix") %>%
        as_tibble(rownames = "OTU")
    # Now join them all into one
    tb <- st %>%
        left_join(sam, by = "Sample") %>%
        left_join(tax, by = "OTU")
    tb
}

```


### load data
```{r, include=FALSE} 
## Phyloseq obj
load("./data/MBLand_ps_land.RData")
ps_land #595 ASVs
load("./data/MBLand_ps_vst.RData")
ps_vst

## DESeq2
load("./data/MBLand_dds_land.RData")
#rowData(dds_land)


## VST
SbySvst <- read.csv( "./data/MBLand_SbyS_vst_2021-11-17.csv", row.names = 1)
dim(SbySvst)
## Euclidean of VST
dist.SbySvst <- stats::dist(SbySvst, method="euclidean")
#dist.SbySvst


## SbyE -- this version has final versions of the soil data
SbyE0 <- read.csv("./data/MBLand_SbyE_2022-01-10.csv", 
                 row.names = 1, stringsAsFactors = TRUE)
dim(SbyE0)
# for now, remove extraneous columns
SbyE <- SbyE0 %>% dplyr::select(-Site.Sample.Order, -Field.Location, -site.rep, 
    -OVT.Var.Entry, -variety.rep, -OVT.plot, -temp, -humid, -leaftemp.diff, -NPQt,
    -Phi2, -PhiNO, -PhiNPQ, -PAR, -Chla, -leafthick, -LAI.ratio.type, -LAI.sd, -LAI.reps,
    LAI.CV, -Notes)
rownames(SbyE) <- SbyE$unique.plot
SbyE$unique.plot <- as.factor(SbyE$unique.plot)
dim(SbyE)


sample_data(ps_land)$nestSite <- as.factor(paste(
        sample_data(ps_land)$species, 
        sample_data(ps_land)$site, sep ="-"))

```

```{r}
taxa <- as.data.frame(as(tax_table(ps_land), "matrix"))
taxa$ASV <- rownames(taxa)

```

# FunGUILD
```{r}
taxa %>% dplyr::rename(Kingdom = domain, Phylum = phylum, Class = class, Order = order, Family = family, Genus = genus, Species = species, OTU_ID = ASV) %>% dplyr::relocate(OTU_ID) -> taxa_table

#taxa_table$Species

#write.table(taxa_table, "./data/MBLand_taxa_table.txt", row.names = FALSE, quote = FALSE)
# NOTE - MASSIVELY MANUALLY EDITED, converted to excel with color coding, re-saved as MBLand_taxa.taxa.txt to match funguild naming convention, just in case
```

```{r}
#guild parser code
#python3 FUNGuild.py guild -taxa example/MBLand_taxa.taxa.txt


# results
FG_results <- read.table("./data/MBLand_taxa.taxa.guilds.txt", 
                    sep = "\t", header = TRUE, stringsAsFactors = TRUE)
#str(FG_results)

table(FG_results$confidenceRanking) 
#only 22 highly probable, 294 are probable
table(FG_results$trophicMode[FG_results$confidenceRanking=='Highly Probable']) 
#highly probable: 9 Symbiotrophs, 10 Saprotrophs, 3 Patho-Sapro
table(FG_results$trophicMode[FG_results$confidenceRanking=='Probable']) 
#probable: 98 Pathotroph, 86 Saprotroph, 3 Symbiotroph, 
#probable:(mixed) 85 Pathotroph-Saprotroph, 4 Pathotroph-Symbiotroph, 11 Saprotroph-Symbiotroph, 7 Pathotroph-Saprotroph-Symbiotroph (catchall)


FG_results %>% 
    filter(confidenceRanking == 'Highly Probable' | confidenceRanking == 'Probable') %>% 
    filter(trophicMode == "Pathotroph" | trophicMode == "Saprotroph" | trophicMode == "Symbiotroph" |
           trophicMode == "Pathotroph-Saprotroph" | 
           trophicMode == "Pathotroph-Symbiotroph" | 
           trophicMode == "Saprotroph-Symbiotroph") -> FG_sub
dim(FG_results); dim(FG_sub)

```

* Out of 595 ASVs, 206 have either highly probable or probable confident trophicModes as one thing only: either patho, sapro, or symbio troph. 309 ASVs have either single or mixed triphicMode categoeies

```{r}
#subest ps_land to only have highly probable/probable confident at one class of trophicMode 
conf_FG <- droplevels(FG_sub$OTU)

#colnames(ps_land@sam_data)[4]
colnames(ps_land@sam_data)[4] <- "crop"  #change name to avoid issues with tax_table 'Species' later
#colnames(ps_land@sam_data)
my_sub <- subset(otu_table(ps_land), rownames(otu_table(ps_land)) %in% conf_FG)
ps_FG <- merge_phyloseq(my_sub, tax_table(ps_land), sample_data(ps_land), refseq(ps_land))
ps_FG #206 ASVs/taxa #309 ASVs with mixed


#suubset funguild output for use as a tax_table()
FG_sub %>% select(OTU, Kingdom, Phylum, Order, Class, Family, Genus, Species, taxon, trophicMode, confidenceRanking, guild, growthForm) -> FG_sub_sub
rownames(FG_sub_sub) <- FG_sub_sub$OTU
FG_sub_sub %>% select(-OTU) -> FG_sub_sub
colnames(FG_sub_sub)


# create a new phyloseq obj with FUNGuild output as the taxa table
merge_phyloseq(otu_table(ps_FG), sample_data(ps_FG), refseq(ps_FG), tax_table(as.matrix(FG_sub_sub))) -> ps_new

#head(ps_new@tax_table)
```

### visualize
```{r}
dat <- tidy_psmelt(ps_new)
#ggplot(dat, aes(x=trophicMode, y = Abundance, fill=crop)) + geom_violin() + geom_point() + scale_y_continuous(trans='sqrt')
```
```{r}
ggplot(dat, aes(x = trophicMode, y = Abundance)) + geom_bar(stat="identity") +
    theme(axis.text.x=element_text(angle=60,hjust=1))
# most ASVs ID-ed as saprotrophs overall
```
```{r}
ggplot(dat, aes(x = Order, y = Abundance, fill = trophicMode)) + geom_bar(stat="identity") +
    theme(axis.text.x=element_text(angle=60,hjust=1))
# most things with assignment are Dothideomycetes
```
```{r, fig.width = 9, fig.height = 6}
ggplot(dat, aes(x = Family, y = Abundance, fill = trophicMode)) + geom_bar(stat="identity") +
    theme(axis.text.x=element_text(angle=60,hjust=1))
# most things with assignment are Mycosphaerellaceae or Phaeosphaeriaceae or Sporidiobolaceae
```
```{r, fig.width = 10, fig.height = 6}
ggplot(dat, aes(x = crop, y = Abundance, fill = Order)) + geom_bar(stat="identity") +
    facet_grid(.~trophicMode) +
    theme(axis.text.x=element_text(angle=60,hjust=1), legend.position = "bottom") +
    guides(fill=guide_legend(ncol=4))
# switchgrass and corn have the most assignments overall
```
```{r, fig.width = 9}
ggplot(dat, aes(x = crop, y = Abundance, fill = nestSite)) + geom_bar(stat="identity") +
    facet_grid(.~trophicMode) + scale_fill_manual("Site", values = site_color) +
    theme(axis.text.x=element_text(angle=60,hjust=1), legend.position = "bottom") 
# moderate differences across sites within crops

```












