---
title: "Fungal Guild Assignments"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
author: "Briana K. Whitaker"
date: "`r Sys.Date()`"
---
\fontsize{9}{10}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning =FALSE, message=FALSE)
```
---




* Run using `r version[['version.string']] `.

# Objective
This document reports on the **fungal guild assignments** from the NC Microbiome Landscapes Project, performed in the summer of 2019. 




# Load Packages, set path to data
```{r, echo=FALSE, results='hide', include=FALSE} 

x<-c("ggplot2", "phyloseq", "dplyr", "RColorBrewer", "DESeq2", "devtools",
     "reshape2",  "viridis", "tidyr", "tidyverse", "RRPP", "car", "olsrr")
lapply(x, require, character.only = TRUE)

#Load functions
source("./code/multiplot.function.R")
source("./code/tidy_psmelt.R")


#add 'not in' function
`%nin%` = Negate(`%in%`)

#set seed
set.seed(829)


# set ggplot2 theme
theme_set(theme_bw(base_size=16)) 
theme_update(panel.grid.major=element_line(0), panel.grid.minor=element_line(0))

# color palettes
display.brewer.pal(8, "Dark2") #assumes Phal is first in order
brewer.pal(8, "Dark2")
spp_color <- c("#E6AB02", "#D95F02", "#1B9E77", '#7570B3') 
#gold, orange, green, purple

#levels(sample_data(ps_filter)$species)
#"CRN"  "SOY"  "SWCH" "WHT" 
#https://mdigi.tools/color-shades/#a6761d
crn_color <- c('#fed561', '#fdc422', '#dda402', '#9e7501')
soy_color <- c('#fea561', '#fd8122', '#dd6102', '#9e4501')
swc_color <- c('#77e8c6', '#40deaf', '#21bf90', '#178867')
wht_color <- c('#cfcde4', '#8f8ac1', '#555096', '#2f2c53')

site_color <- c(crn_color, soy_color, swc_color, wht_color)

```

```{r, results = 'hide'} 
## Phyloseq obj
load("./data/MBLand_ps_land.RData")
ps_land #595 ASVs
load("./data/MBLand_ps_vst.RData")
ps_vst


sample_data(ps_land)$nestSite <- as.factor(paste(
        sample_data(ps_land)$species, 
        sample_data(ps_land)$site, sep ="-"))


#colnames(ps_land@sam_data)[4]
colnames(ps_land@sam_data)[4] <- "crop"  #change name to avoid issues with tax_table 'Species' later
#colnames(ps_land@sam_data)

## DESeq2
load("./data/MBLand_dds_land.RData")
#rowData(dds_land)


## VST
SbySvst <- read.csv( "./data/MBLand_SbyS_vst_2021-11-17.csv", row.names = 1)
dim(SbySvst)
## Euclidean of VST
dist.SbySvst <- stats::dist(SbySvst, method="euclidean")
#dist.SbySvst


## SbyE -- this version has final versions of the soil data
SbyE0 <- read.csv("./data/MBLand_SbyE_2022-01-10.csv", 
                 row.names = 1, stringsAsFactors = TRUE)
dim(SbyE0)



# site level data

site_dat0 <- read.csv("./data/MBLand_Rich_GEE_means.csv",
                  row.names = 1, stringsAsFactors = TRUE)

```

```{r, include=FALSE}
taxa <- as.data.frame(as(tax_table(ps_land), "matrix"))
taxa$ASV <- rownames(taxa)

```

# FunGUILD
```{r, include = FALSE}
taxa %>% dplyr::rename(Kingdom = domain, Phylum = phylum, Class = class, 
                       Order = order, Family = family, Genus = genus, 
                       Species = species, OTU_ID = ASV) %>% 
    dplyr::relocate(OTU_ID) -> taxa_table

#taxa_table$Species

#write.table(taxa_table, "./data/MBLand_taxa_table.txt", 
#    row.names = FALSE, quote = FALSE)
# #NOTE - Manually edited, converted to excel with color coding, re-saved as 
# #  MBLand_taxa.taxa.txt to match funguild naming convention
```

```{r, include = FALSE}
#guild parser code
#python3 FUNGuild.py guild -taxa example/MBLand_taxa.taxa.txt


# results
FG_results <- read.table("./data/MBLand_taxa.taxa.guilds.txt", 
                    sep = "\t", header = TRUE, stringsAsFactors = TRUE)
#str(FG_results)

table(FG_results$confidenceRanking) 
#only 22 highly probable, 294 are probable
table(FG_results$trophicMode[FG_results$confidenceRanking=='Highly Probable']) 
#highly probable: 9 Symbiotrophs, 10 Saprotrophs, 3 Patho-Sapro
table(FG_results$trophicMode[FG_results$confidenceRanking=='Probable']) 
#probable: 98 Pathotroph, 86 Saprotroph, 3 Symbiotroph, 
#probable:(mixed) 85 Pathotroph-Saprotroph, 4 Pathotroph-Symbiotroph, 11 Saprotroph-Symbiotroph, 7 Pathotroph-Saprotroph-Symbiotroph (catchall)

```


* Out of 595 ASVs, 206 have either highly probable or probable confident trophicModes as one thing only: either patho, sapro, or symbio troph. 309 ASVs have either single or mixed triphicMode categoeies. 316 ASVs have single, double or triple guild assignments



#### Summarize Trophic Modes to Fewer Categories
```{r, results = 'hide', include=FALSE}
FG_results$myTrophicMode <- ifelse(FG_results$confidenceRanking != "Possible", 
            as.character(FG_results$trophicMode), 'na')
FG_results$myTrophicMode <- factor(FG_results$myTrophicMode)
#table(FG_results$myTrophicMode) #202+77=279 na's
levels(FG_results$myTrophicMode) <- list(
    "Pathotroph" = c("Pathotroph"),
    "Symbiotroph" = c("Symbiotroph"),
    "Saprotroph" = c("Saprotroph"),
    "na" = c("Saprotroph-Symbiotroph", "Pathotroph-Saprotroph", "Pathotroph-Symbiotroph",
             "na", "Pathotroph-Saprotroph-Symbiotroph"))
table(FG_results$myTrophicMode)


#subset funguild output for use as a tax_table()
FG_results %>% select(OTU, Kingdom, Phylum, Order, Class, Family, 
                      Genus, Species, taxon, trophicMode, 
                      confidenceRanking, myTrophicMode) -> FG_sub_all
rownames(FG_sub_all) <- FG_sub_all$OTU
FG_sub_all %>% select(-OTU) -> FG_sub_all
colnames(FG_sub_all)


# create a new phyloseq obj with FUNGuild output as the taxa table
merge_phyloseq(otu_table(ps_land), sample_data(ps_land),
               refseq(ps_land), tax_table(as.matrix(FG_sub_all))) -> ps_FG_all
#head(ps_FG_all@tax_table)

# tidy melt
FG_melt <- tidy_psmelt(ps_FG_all)
#formerly dat
#dim(FG_melt)  #113645=191*595 
#dim(FG_results)
 
```

#### Plot summarized trophic modes
```{r}
#make variables into factors after melt process
FG_melt$myTrophicMode <- factor(FG_melt$myTrophicMode)

ggplot(FG_melt, aes(x = crop, y = Abundance, fill = nestSite)) +
    geom_bar(stat="identity") +
    facet_grid(.~myTrophicMode) + 
    scale_fill_manual("Site", values = site_color) +
    theme(axis.text.x=element_text(angle=60,hjust=1), 
          legend.position = "bottom")
```

* Most of the assignments are in the NA category. Very few pathogens or symbionts.

#### Per Sample Abundance calculation
```{r, include=FALSE}
# uniList <- levels(FG_melt$unique.plot)
# trophList <- levels(FG_melt$myTrophicMode)
# out <- list()
# for (i in (1:length(uniList))) {
#   myPlot <- FG_melt$unique.plot[i]
#   allSum <- list()
#   #print(as.character(myPlot))                         #debug line
#   for (j in (1:length(trophList))) {
#     myTroph <- levels(FG_melt$myTrophicMode)[j]
#     #print(as.character(myTroph))                      #debug line
# 
#     #start
#     require(dplyr)
#     fg <- FG_melt %>%
#         filter(myTrophicMode == myTroph) %>%
#         filter(unique.plot == myPlot)
#     mySum <- sum(fg$Abundance)
#     allSum <- base::rbind(allSum,
#                          cbind("plot" = as.character(myPlot),
#                                "troph" = as.character(myTroph),
#                                "sum" = mySum))
#     #print(allSum)
#   }
#   rm(fg)                                              #cleanup
#   out <- base::rbind(out, allSum)
#   rm(allSum)                                          #cleanup
#   #print(out)                                          #debug line
#   }
# dim(out)
# rm(i, j, mySum, myPlot, myTroph)              #cleanup
# #write.csv(out, "./data/MBLand_FunGuild_TrophicAbundance.csv")

```

* Organize and rearrange. Convert data to proportional variables, relative to sequencing depth per sample. (code hidden)

```{r, include = FALSE}
FG_sumAbund <- read.csv("./data/MBLand_FunGuild_TrophicAbundance.csv", 
         header = TRUE, colClasses = c('character', 'character', 
                                       'factor', 'character')) #otherwise does not read plot correctly
FG_sumAbund$sum <- as.numeric(FG_sumAbund$sum)
FG_sumAbund$plot <- as.factor(FG_sumAbund$plot)

# pivot-wider
SbyE_fg <- pivot_wider(FG_sumAbund, id_cols = "plot", 
                       names_from = c("troph"), values_from = "sum")
SbyE_fg <- as.data.frame(SbyE_fg)
rownames(SbyE_fg) <- SbyE_fg$plot

# creaete new and merge phyloseq
ps_new <- sample_data(SbyE_fg)
ps_FG <- merge_phyloseq(ps_land, ps_new)

# create proportional variables, relative to seq depth
sample_data(ps_FG)$prop_Pathotroph <- 
    c(sample_data(ps_FG)$Pathotroph/sample_data(ps_FG)$seqDepth_onlyFungi)
sample_data(ps_FG)$prop_Saprotroph <- 
    c(sample_data(ps_FG)$Saprotroph/sample_data(ps_FG)$seqDepth_onlyFungi)
sample_data(ps_FG)$prop_Symbiotroph <-
    c(sample_data(ps_FG)$Symbiotroph/sample_data(ps_FG)$seqDepth_onlyFungi)
sample_data(ps_FG)$prop_na <- 
    c(sample_data(ps_FG)$na/sample_data(ps_FG)$seqDepth_onlyFungi)

# #visualize
# plot(log2(ps_FG@sam_data$prop_Pathotroph) ~ log2(ps_FG@sam_data$prop_Saprotroph))
# plot(log2(ps_FG@sam_data$prop_Pathotroph) ~ log2(ps_FG@sam_data$prop_Symbiotroph))
# plot(log2(ps_FG@sam_data$prop_Pathotroph) ~ log2(ps_FG@sam_data$prop_na))
# plot(log2(ps_FG@sam_data$prop_Saprotroph) ~ log2(ps_FG@sam_data$prop_Symbiotroph))
# plot(log2(ps_FG@sam_data$prop_Saprotroph) ~ log2(ps_FG@sam_data$prop_na))
# plot(log2(ps_FG@sam_data$prop_Symbiotroph) ~ log2(ps_FG@sam_data$prop_na))

#organize for linear models
SbyE_new <- data.frame(sample_data(ps_FG)@.Data)
colnames(SbyE_new) <- ps_FG@sam_data@names
rownames(SbyE_new) <- ps_FG@sam_data@row.names

```


# RRPP - Do proportion pathos differ by crop and site?
* Does not include Variety (would need to split by crop)
* Performed at the plot level

```{r, fig.width = 7, include = FALSE}
par(mfrow = c(1,1))
#hist(log(I(SbyE_new$prop_Pathotroph+0.0001)))
# alternative
#hist(asin(sqrt(SbyE_new$prop_Pathotroph)), breaks = 20)


#hist(log(I(SbyE_new$prop_Symbiotroph+0.0001)))
#hist(sqrt(SbyE_new$prop_Saprotroph))
# should be transformed if run in a typical linear model. Likely fine for RRPP/permutation procedure
par(mfrow = c(1,1))
```

```{r, results = 'hide'}
fg_mod <- lm.rrpp(prop_Pathotroph ~ 
                          crop +
                          crop:site +
                          crop:site:plot.rep, iter = 999,
                          print.progress = FALSE, data = SbyE_new, SS.type="III")
```

```{r}
full.anova <- anova(fg_mod, effect.type = "F", error = c("crop:site:plot.rep",
    "crop:site:plot.rep", "Residuals"))
summary(full.anova, formula = false)
```

```{r, results = 'hide', echo = FALSE, fig.width = 5}
path_bySite <- ggplot(SbyE_new, aes(y = I((prop_Pathotroph+0.0001)^(1/4)), x = nestSite, fill = factor(nestSite))) +
    geom_boxplot(outlier.shape = NULL) + scale_fill_manual(values = site_color) +
    geom_jitter(width = .1)+
    scale_x_discrete("Site") +
    scale_y_continuous("Proportion of Pathogen Sequences\n(transformed)",
                       limits = c(0.09,0.9)) + 
    #coord_trans(y= scales::boxcox_trans(0.25, offset = 0.01)) +
    guides(fill = 'none') +
    theme(axis.text.x = element_text(angle = 90))

path_bySite

```

#### CVs of Proportions across Sites
```{r}
patho_CV <- c(round(tapply(SbyE_new$prop_Pathotroph, SbyE_new$nestSite,
    stats::sd)/tapply(SbyE_new$prop_Pathotroph, SbyE_new$nestSite, mean),4)*100)

data.frame(patho_CV)

```

* These C.V.'s are quite high, above 100% in some cases for pathogen . 

# Best Subsets - Proportion Pathogens 
* Performed at the site level.
* No soil variables, as with the Richness and Hetoerogeneity models

```{r, include = FALSE}
out <- data.frame(
    nestSite=c(levels(SbyE_new$nestSite)),
    #crop = c(rep(levels(SbyE_new$crop), each = 4)),

    prop_Pathotroph=as.vector(tapply(SbyE_new$prop_Pathotroph, 
              list(SbyE_new$nestSite), mean)),
    prop_Pathotroph_sd=as.vector(tapply(SbyE_new$prop_Pathotroph, 
              list(SbyE_new$nestSite), sd)),
    prop_Pathotroph_reps=as.vector(table(SbyE_new$nestSite)) )
out$prop_Pathotroph_se <- out$prop_Pathotroph_sd/sqrt(out$prop_Pathotroph_reps)
out %>% dplyr::select(-prop_Pathotroph_reps, -prop_Pathotroph_sd) -> out

site_dat <- merge(site_dat0, out, by.x = "site", by.y = "nestSite")

#hist(log(site_dat$prop_Pathotroph))  #log transformations work for both vars
#hist(log(site_dat$prop_Pathotroph_se))

```


```{r}
patho1 <- lm(log(prop_Pathotroph) ~  #Lat.N + Lon.W + Elev +  MLT + MHT +  
                                     #needs to be removed to run, matches CCA/richness models
                 MAP + Precip_90day + LT_90day + HT_90day + 
                 Perc_crop_10km + Perc_veg_10km + Perc_crop_1km + Perc_veg_1km +
                 soilmoist + LAI.mean + plantHt, data=site_dat)
Anova(patho1)
olsrr::ols_plot_resid_fit(patho1) 

olsrr::ols_test_normality(patho1)

olsrr::ols_test_breusch_pagan(patho1, rhs=TRUE, multiple=TRUE) 

olsrr::ols_coll_diag(patho1)$vif_t

```

```{r}
patho2 <- lm(log(prop_Pathotroph) ~  #HT_90day + 
                 MAP + Precip_90day + LT_90day + 
                 Perc_crop_10km + Perc_veg_10km + Perc_crop_1km + Perc_veg_1km +
                 soilmoist + LAI.mean + plantHt, data=site_dat)
olsrr::ols_coll_diag(patho2)$vif_t

patho3 <- lm(log(prop_Pathotroph) ~  #HT_90day + Perc_veg_10km + 
                 MAP + Precip_90day + LT_90day + 
                 Perc_crop_10km + Perc_crop_1km + Perc_veg_1km +
                 soilmoist + LAI.mean + plantHt, data=site_dat)
olsrr::ols_coll_diag(patho3)$vif_t

patho4 <- lm(log(prop_Pathotroph) ~  #HT_90day + Perc_veg_10km + Perc_crop_10km + 
                 MAP + Precip_90day + LT_90day + 
                 Perc_crop_1km + Perc_veg_1km +
                 soilmoist + LAI.mean + plantHt, data=site_dat)
olsrr::ols_coll_diag(patho4)$vif_t

patho5 <- lm(log(prop_Pathotroph) ~  #HT_90day + Perc_veg_10km + Perc_crop_10km + 
                                     #Perc_crop_1km + 
                 MAP + Precip_90day + LT_90day + 
                 Perc_veg_1km +
                 soilmoist + LAI.mean + plantHt, data=site_dat)
olsrr::ols_coll_diag(patho5)$vif_t

patho6 <- lm(log(prop_Pathotroph) ~  #HT_90day + Perc_veg_10km + Perc_crop_10km + 
                                     #Perc_crop_1km + plantHt + 
                 MAP + Precip_90day + LT_90day + 
                 Perc_veg_1km +
                 soilmoist + LAI.mean, data=site_dat)
olsrr::ols_coll_diag(patho6)$vif_t

```

```{r}
#olsrr::ols_plot_resid_fit(patho6)  #looks great
olsrr::ols_step_best_subset(patho6)

patho6$coefficients
```
* Model 2 lowest SBC, Model 3 lowest AIC, Model 1 lowest SBIC. Model 3 highest Adj.R2. Models 2-6 essentially plateau Adj.R2 (model 1 much less). 
* Model 3 = LT_90day, Perc_veg_1km, LAI.mean




# Best Subsets - Standard-Error Pathogens 
* Performed at the site level.
* No soil variables, as with the Richness and Hetoerogeneity models

```{r}
se1 <- lm(log(prop_Pathotroph_se) ~  #Lat.N + Lon.W + Elev +  MLT + MHT +  
                                     #needs to be removed to run, matches CCA/richness models
                 MAP + Precip_90day + LT_90day + HT_90day + 
                 Perc_crop_10km + Perc_veg_10km + Perc_crop_1km + Perc_veg_1km +
                 soilmoist + LAI.mean + plantHt, data=site_dat)
Anova(se1)
olsrr::ols_plot_resid_fit(se1) 

olsrr::ols_test_normality(se1)

olsrr::ols_test_breusch_pagan(se1, rhs=TRUE, multiple=TRUE) 

olsrr::ols_coll_diag(se1)$vif_t

```

```{r}
se2 <- lm(log(prop_Pathotroph_se) ~  #HT_90day + 
                 MAP + Precip_90day + LT_90day + 
                 Perc_crop_10km + Perc_veg_10km + Perc_crop_1km + Perc_veg_1km +
                 soilmoist + LAI.mean + plantHt, data=site_dat)
olsrr::ols_coll_diag(se2)$vif_t

se3 <- lm(log(prop_Pathotroph_se) ~  #HT_90day + Perc_veg_10km + 
                 MAP + Precip_90day + LT_90day + 
                 Perc_crop_10km + Perc_crop_1km + Perc_veg_1km +
                 soilmoist + LAI.mean + plantHt, data=site_dat)
olsrr::ols_coll_diag(se3)$vif_t

se4 <- lm(log(prop_Pathotroph_se) ~  #HT_90day + Perc_veg_10km + Perc_crop_10km + 
                 MAP + Precip_90day + LT_90day + 
                 Perc_crop_1km + Perc_veg_1km +
                 soilmoist + LAI.mean + plantHt, data=site_dat)
olsrr::ols_coll_diag(se4)$vif_t

se5 <- lm(log(prop_Pathotroph_se) ~  #HT_90day + Perc_veg_10km + Perc_crop_10km + 
                                     #Perc_crop_1km + 
                 MAP + Precip_90day + LT_90day + 
                 Perc_veg_1km +
                 soilmoist + LAI.mean + plantHt, data=site_dat)
olsrr::ols_coll_diag(se5)$vif_t

se6 <- lm(log(prop_Pathotroph_se) ~  #HT_90day + Perc_veg_10km + Perc_crop_10km + 
                                     #Perc_crop_1km + plantHt 
                 MAP + Precip_90day + LT_90day + 
                 Perc_veg_1km +
                 soilmoist + LAI.mean, data=site_dat)
olsrr::ols_coll_diag(se6)$vif_t
```

```{r}
#olsrr::ols_plot_resid_fit(se6)  #looks great
olsrr::ols_step_best_subset(se6)

se6$coefficients

```
* Model 3 likely the best, highest AdjR2 (59%!), lowest AIC and SBC, 3rd lowest SBIC
* Model 3 = LT_90day, Perc_veg_1km, LAI.mean



# FIGURES
```{r}

p1 <- ggplot(site_dat,
       aes(x = Perc_veg_1km, y = prop_Pathotroph, colour = site)) +
  geom_smooth(aes(x = Perc_veg_1km, y = prop_Pathotroph), 
              method=lm, formula = y ~ x,
              se = TRUE, color = '#666666', linetype = 1) +
  geom_point(size = 3) +
  scale_y_continuous("Proportion Pathotroph", trans = "log", 
                     breaks = c(0.02, 0.05, 0.1, 0.2, 0.5), limits = c(0.002,1.2)) +
  scale_x_continuous("% Unmanaged Vegetation\n1-km scale", limits = c(0,55)) +
  scale_color_manual("", values = site_color) +
  theme(legend.position="none") +
  annotate(geom = "text", x = 3, y = 1, label = "A)", size = 7)
p2 <- ggplot(site_dat,
       aes(x = LT_90day, y = prop_Pathotroph, colour = site)) +
  geom_smooth(aes(x = LT_90day, y = prop_Pathotroph), 
              method=lm, formula = y ~ x,
              se = TRUE, color = '#666666', linetype = 1) +
  geom_point(size = 3) +
  scale_y_continuous("Proportion Pathotroph", trans = "log", 
                     breaks = c(0.02, 0.05, 0.1, 0.2, 0.5), limits = c(0.002,1.2)) +
  scale_x_continuous(expression(atop("90-day Low Temperature", 
                            paste("(",~degree, "C)", sep = ""))), 
                                limits = c(-5.5,11)) +
  scale_color_manual("", values = site_color) +
  theme(legend.position="none") +
  annotate(geom = "text", x = -4.7, y = 1, label = "B)", size = 7)
p3 <- ggplot(site_dat,
       aes(x = LAI.mean, y = prop_Pathotroph, colour = site)) +
  geom_smooth(aes(x = LAI.mean, y = prop_Pathotroph), 
              method=lm, formula = y ~ x,
              se = TRUE, color = '#666666', linetype = 1) +
  geom_point(size = 3) +
  scale_y_continuous("Proportion Pathotroph", trans = "log", 
                     breaks = c(0.02, 0.05, 0.1, 0.2, 0.5), limits = c(0.002,1.2)) +
  scale_x_continuous("Leaf Area Index", limits = c(3.2,12.5)) +
  scale_color_manual("", values = site_color) +
  theme(legend.position="none") +
  annotate(geom = "text", x = 3.65, y = 1, label = "C)", size = 7)

#multiplot(p1, p2, p3, cols =3)

p4 <- ggplot(site_dat,
       aes(x = Perc_veg_1km, y = prop_Pathotroph_se, colour = site)) +
  geom_smooth(aes(x = Perc_veg_1km, y = prop_Pathotroph_se), 
              method=lm, formula = y ~ x,
              se = TRUE, color = '#666666', linetype = 1) +
  geom_point(size = 3) +
  scale_y_continuous("St. Error of Proportion Pathotroph", trans = "log", 
                     breaks = c(0.002, 0.005, 0.01, 0.02, 0.05), limits = c(0.0006,0.1)) + 
  scale_x_continuous("% Unmanaged Vegetation\n1-km scale", limits = c(0,55)) +
  scale_color_manual("", values = site_color) +
  theme(legend.position="none") +
  annotate(geom = "text", x = 3, y = 0.083, label = "D)", size = 7)
p5 <- ggplot(site_dat,
       aes(x = LT_90day, y = prop_Pathotroph_se, colour = site)) +
  geom_smooth(aes(x = LT_90day, y = prop_Pathotroph_se), 
              method=lm, formula = y ~ x,
              se = TRUE, color = '#666666', linetype = 1) +
  geom_point(size = 3) +
  scale_y_continuous("St. Error of Proportion Pathotroph", trans = "log", 
                     breaks = c(0.002, 0.005, 0.01, 0.02, 0.05), limits = c(0.0006,0.1)) +
  scale_x_continuous(expression(atop("90-day Low Temperature", 
                            paste("(",~degree, "C)", sep = ""))), 
                                limits = c(-5.5,11)) +
  scale_color_manual("", values = site_color) +
  theme(legend.position="none") +
  annotate(geom = "text", x = -4.7, y = 0.083, label = "E)", size = 7)
p6 <- ggplot(site_dat,
       aes(x = LAI.mean, y = prop_Pathotroph_se, colour = site)) +
  geom_smooth(aes(x = LAI.mean, y = prop_Pathotroph_se), 
              method=lm, formula = y ~ x,
              se = TRUE, color = '#666666', linetype = 1) +
  geom_point(size = 3) +
  scale_y_continuous("St. Error of Proportion Pathotroph", trans = "log", 
                     breaks = c(0.002, 0.005, 0.01, 0.02, 0.05), limits = c(0.0006,0.1)) +
  scale_x_continuous("Leaf Area Index", limits = c(3.2,12.5)) +
  scale_color_manual("", values = site_color) +
  theme(legend.position="none") +
  annotate(geom = "text", x = 3.65, y = 0.083, label = "F)", size = 7)

#tiff("./figures/Prop Pathotroph ~ Vars.tiff", width=13.2, height=9, units="in", res=600)
multiplot(p1, p2, p3, p4, p5, p6, cols = 3)
#dev.off()
# tiff("test.tiff", width=13.2, height=9, units="in", res=600)
# p1 + guides(colour=guide_legend(ncol=8)) + theme(legend.position = "bottom")
# dev.off()
```



# SUMMARY
* So when it is more recently warm, the plants are bigger, and there is more native vegetation, there are both more pathogens within fields and more variable pathogens abundance across plots within fields?




# CORRELATIONS
```{r}
ggplot(site_dat, 
       aes(x=Rich_nosing, y=prop_Pathotroph)) +
    geom_point(aes(colour=site), size = 3) +
    scale_colour_manual(values=site_color) +
    scale_x_continuous("Avg. Within Site Richness") +
    scale_y_continuous("Avg. Within Site Pathogen Abundance", trans = "log", 
                       breaks = c(0.02, 0.05, 0.1, 0.2, 0.5))

ggplot(site_dat, 
       aes(x=mean, y=prop_Pathotroph_se)) +
    geom_point(aes(colour=site), size = 3) +
    scale_colour_manual(values=site_color) +
    scale_x_continuous("Avg. Within Site Dissimilarity") +
    scale_y_continuous("Avg. Within Site Pathogen SE", trans = "log", 
                       breaks = c(0.002, 0.005, 0.01, 0.02, 0.05))


cor(site_dat[,c(3,28,37,38)])


```



#### end
```{r}
sessionInfo()
```

