---
title: "Fungal Guild Assignments"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
author: "Briana K. Whitaker"
date: "`r Sys.Date()`"
---
\fontsize{9}{10}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning =FALSE, message=FALSE)
```
---




* Run using `r version[['version.string']] `.

# Objective
This document reports on the **fungal guild assignments** from the NC Microbiome Landscapes Project, performed in the summer of 2019. 




# Load Packages, set path to data
```{r, echo=FALSE, results='hide', include=FALSE} 

x<-c("ggplot2", "phyloseq", "dplyr", "RColorBrewer", "DESeq2", "devtools",
     "reshape2",  "viridis", "tidyr", "tidyverse", "RRPP", "car")
lapply(x, require, character.only = TRUE)

#Load functions
source("./code/multiplot.function.R")
source("./code/tidy_psmelt.R")


#add 'not in' function
`%nin%` = Negate(`%in%`)

#set seed
set.seed(829)


# set ggplot2 theme
theme_set(theme_bw(base_size=16)) 
theme_update(panel.grid.major=element_line(0), panel.grid.minor=element_line(0))

# color palettes
display.brewer.pal(8, "Dark2") #assumes Phal is first in order
brewer.pal(8, "Dark2")
spp_color <- c("#E6AB02", "#D95F02", "#1B9E77", '#7570B3') 
#gold, orange, green, purple

#levels(sample_data(ps_filter)$species)
#"CRN"  "SOY"  "SWCH" "WHT" 
#https://mdigi.tools/color-shades/#a6761d
crn_color <- c('#fed561', '#fdc422', '#dda402', '#9e7501')
soy_color <- c('#fea561', '#fd8122', '#dd6102', '#9e4501')
swc_color <- c('#77e8c6', '#40deaf', '#21bf90', '#178867')
wht_color <- c('#cfcde4', '#8f8ac1', '#555096', '#2f2c53')

site_color <- c(crn_color, soy_color, swc_color, wht_color)

```

```{r, results = 'hide'} 
## Phyloseq obj
load("./data/MBLand_ps_land.RData")
ps_land #595 ASVs
load("./data/MBLand_ps_vst.RData")
ps_vst

## DESeq2
load("./data/MBLand_dds_land.RData")
#rowData(dds_land)


## VST
SbySvst <- read.csv( "./data/MBLand_SbyS_vst_2021-11-17.csv", row.names = 1)
dim(SbySvst)
## Euclidean of VST
dist.SbySvst <- stats::dist(SbySvst, method="euclidean")
#dist.SbySvst


## SbyE -- this version has final versions of the soil data
SbyE0 <- read.csv("./data/MBLand_SbyE_2022-01-10.csv", 
                 row.names = 1, stringsAsFactors = TRUE)
dim(SbyE0)
```

```{r, include=FALSE}
# for now, remove extraneous columns
SbyE <- SbyE0 %>% dplyr::select(-Site.Sample.Order, -Field.Location, -site.rep, 
    -OVT.Var.Entry, -variety.rep, -OVT.plot, -temp, -humid, -leaftemp.diff, -NPQt,
    -Phi2, -PhiNO, -PhiNPQ, -PAR, -Chla, -leafthick, -LAI.ratio.type, -LAI.sd, -LAI.reps,
    LAI.CV, -Notes)
rownames(SbyE) <- SbyE$unique.plot
SbyE$unique.plot <- as.factor(SbyE$unique.plot)
dim(SbyE)


sample_data(ps_land)$nestSite <- as.factor(paste(
        sample_data(ps_land)$species, 
        sample_data(ps_land)$site, sep ="-"))

```

```{r}
taxa <- as.data.frame(as(tax_table(ps_land), "matrix"))
taxa$ASV <- rownames(taxa)

```

# FunGUILD
```{r, include = FALSE}
taxa %>% dplyr::rename(Kingdom = domain, Phylum = phylum, Class = class, 
                       Order = order, Family = family, Genus = genus, 
                       Species = species, OTU_ID = ASV) %>% 
    dplyr::relocate(OTU_ID) -> taxa_table

#taxa_table$Species

#write.table(taxa_table, "./data/MBLand_taxa_table.txt", 
#    row.names = FALSE, quote = FALSE)
# #NOTE - Manually edited, converted to excel with color coding, re-saved as 
# #  MBLand_taxa.taxa.txt to match funguild naming convention
```

```{r}
#guild parser code
#python3 FUNGuild.py guild -taxa example/MBLand_taxa.taxa.txt


# results
FG_results <- read.table("./data/MBLand_taxa.taxa.guilds.txt", 
                    sep = "\t", header = TRUE, stringsAsFactors = TRUE)
#str(FG_results)

table(FG_results$confidenceRanking) 
#only 22 highly probable, 294 are probable
table(FG_results$trophicMode[FG_results$confidenceRanking=='Highly Probable']) 
#highly probable: 9 Symbiotrophs, 10 Saprotrophs, 3 Patho-Sapro
table(FG_results$trophicMode[FG_results$confidenceRanking=='Probable']) 
#probable: 98 Pathotroph, 86 Saprotroph, 3 Symbiotroph, 
#probable:(mixed) 85 Pathotroph-Saprotroph, 4 Pathotroph-Symbiotroph, 11 Saprotroph-Symbiotroph, 7 Pathotroph-Saprotroph-Symbiotroph (catchall)
```

### Visualization - Original Data
```{r, results = 'hide'}
FG_results %>% 
    filter(confidenceRanking == 'Highly Probable' | 
           confidenceRanking == 'Probable') %>%
    filter(trophicMode != "Pathotroph-Saprotroph-Symbiotroph") -> FG_sub         #%>% 
    # filter(trophicMode == "Pathotroph" | 
    #        trophicMode == "Saprotroph" | 
    #        trophicMode == "Symbiotroph" |
    #        trophicMode == "Pathotroph-Saprotroph" | 
    #        trophicMode == "Pathotroph-Symbiotroph" | 
    #        trophicMode == "Saprotroph-Symbiotroph") -> FG_sub
dim(FG_results); dim(FG_sub)

```

* Out of 595 ASVs, 206 have either highly probable or probable confident trophicModes as one thing only: either patho, sapro, or symbio troph. 309 ASVs have either single or mixed triphicMode categoeies. 316 ASVs have single, double or triple guild assignments

```{r, results = 'hide'}
#subest ps_land to only have highly probable/probable confident
conf_FG <- droplevels(FG_sub$OTU)

#colnames(ps_land@sam_data)[4]
colnames(ps_land@sam_data)[4] <- "crop"  #change name to avoid issues with tax_table 'Species' later
#colnames(ps_land@sam_data)
my_sub <- subset(otu_table(ps_land), rownames(otu_table(ps_land)) %in% conf_FG)
ps_FG <- merge_phyloseq(my_sub, tax_table(ps_land), 
                        sample_data(ps_land), refseq(ps_land))
ps_FG #309 ASVs


#subset funguild output for use as a tax_table()
FG_sub %>% select(OTU, Kingdom, Phylum, Order, Class, Family, Genus, 
                  Species, taxon, trophicMode, confidenceRanking, 
                  guild, growthForm) -> FG_sub_sub
rownames(FG_sub_sub) <- FG_sub_sub$OTU
FG_sub_sub %>% select(-OTU) -> FG_sub_sub
colnames(FG_sub_sub)


# create a new phyloseq obj with FUNGuild output as the taxa table
merge_phyloseq(otu_table(ps_FG), sample_data(ps_FG), refseq(ps_FG), 
               tax_table(as.matrix(FG_sub_sub))) -> ps_new

#head(ps_new@tax_table)

# tidy melt
dat <- tidy_psmelt(ps_new)
rm(ps_new)
rm(ps_FG)

```

```{r, echo = FALSE}
ggplot(dat, aes(x=trophicMode, y = Abundance, fill=crop)) + 
    geom_bar(stat="identity") + scale_y_continuous(trans='sqrt') +
    theme(axis.text.x=element_text(angle=60,hjust=1)) +
    scale_fill_manual("Site", values = spp_color)
# most ASVs ID-ed as saprotrophs overall
```
```{r, echo = FALSE}
ggplot(dat, aes(x = Order, y = Abundance, fill = trophicMode)) + geom_bar(stat="identity") +
    theme(axis.text.x=element_text(angle=60,hjust=1))
# most things with assignment are Dothideomycetes
```
```{r, echo = FALSE, fig.width = 9, fig.height = 6}
# ggplot(dat, aes(x = Family, y = Abundance, fill = trophicMode)) + geom_bar(stat="identity") +
#     theme(axis.text.x=element_text(angle=60,hjust=1))
# # most things with assignment are Mycosphaerellaceae or Phaeosphaeriaceae or Sporidiobolaceae
```
```{r, echo = FALSE, fig.width = 10, fig.height = 6}
# ggplot(dat, aes(x = crop, y = Abundance, fill = Order)) + geom_bar(stat="identity") +
#     facet_grid(.~trophicMode) +
#     theme(axis.text.x=element_text(angle=60,hjust=1), legend.position = "bottom") +
#     guides(fill=guide_legend(ncol=4))
# # switchgrass and corn have the most assignments overall
```



# Summarize Trophic Modes to Fewer Categories
```{r, results = 'hide'}
FG_results$myTrophicMode <- ifelse(FG_results$confidenceRanking != "Possible", 
            as.character(FG_results$trophicMode), 'na')
FG_results$myTrophicMode <- factor(FG_results$myTrophicMode)
#table(FG_results$myTrophicMode) #202+77=279 na's
levels(FG_results$myTrophicMode) <- list(
    "Patho_other" = c("Pathotroph", "Pathotroph-Saprotroph"),
    "Symbio_other" = c("Symbiotroph", "Saprotroph-Symbiotroph"),
    "Other" = c("Saprotroph", "Pathotroph-Symbiotroph"),
    "na" = c("na", "Pathotroph-Saprotroph-Symbiotroph"))
table(FG_results$myTrophicMode)


#subset funguild output for use as a tax_table()
FG_results %>% select(OTU, Kingdom, Phylum, Order, Class, Family, 
                      Genus, Species, taxon, trophicMode, 
                      confidenceRanking, myTrophicMode) -> FG_sub_all
rownames(FG_sub_all) <- FG_sub_all$OTU
FG_sub_all %>% select(-OTU) -> FG_sub_all
colnames(FG_sub_all)


# create a new phyloseq obj with FUNGuild output as the taxa table
merge_phyloseq(otu_table(ps_land), sample_data(ps_land),
               refseq(ps_land), tax_table(as.matrix(FG_sub_all))) -> ps_FG_all
#head(ps_FG_all@tax_table)

# tidy melt
FG_melt <- tidy_psmelt(ps_FG_all)
#formerly dat
#dim(FG_melt)  #113645=191*595 
#dim(FG_results)
 
```

### Plot summarized trophic modes
```{r}
#make variables into factors after melt process
FG_melt$myTrophicMode <- factor(FG_melt$myTrophicMode)

ggplot(FG_melt, aes(x = crop, y = Abundance, fill = nestSite)) +
    geom_bar(stat="identity") +
    facet_grid(.~myTrophicMode) + 
    scale_fill_manual("Site", values = site_color) +
    theme(axis.text.x=element_text(angle=60,hjust=1), 
          legend.position = "bottom")
```

* Most of the assignments are in the NA or Other category. Very few pathogens or symbionts.

### Per Sample Abundance calculation
```{r}
# uniList <- levels(FG_melt$unique.plot)
# trophList <- levels(FG_melt$myTrophicMode)
# out <- list()
# for (i in (1:length(uniList))) {
#   myPlot <- FG_melt$unique.plot[i]
#   allSum <- list()
#   #print(as.character(myPlot))                         #debug line
#   for (j in (1:length(trophList))) {
#     myTroph <- levels(FG_melt$myTrophicMode)[j]
#     #print(as.character(myTroph))                      #debug line
# 
#     #start
#     require(dplyr)
#     fg <- FG_melt %>%
#         filter(myTrophicMode == myTroph) %>%
#         filter(unique.plot == myPlot)
#     mySum <- sum(fg$Abundance)
#     allSum <- base::rbind(allSum,
#                          cbind("plot" = as.character(myPlot),
#                                "troph" = as.character(myTroph),
#                                "sum" = mySum))
#     #print(allSum)
#   }
#   rm(fg)                                              #cleanup
#   out <- base::rbind(out, allSum)
#   rm(allSum)                                          #cleanup
#   #print(out)                                          #debug line
#   }
# dim(out)
# rm(i, j, mySum, myPlot, myTroph)              #cleanup
# #write.csv(out, "./data/MBLand_FunGuild_TrophicAbundance.csv")
```

* Organize and rearrange. Convert data to proportional variables, relative to sequencing depth per sample. (code hidden)

```{r, include = FALSE}
FG_sumAbund <- read.csv("./data/MBLand_FunGuild_TrophicAbundance.csv", 
         header = TRUE, colClasses = c('character', 'character', 
                                       'factor', 'character')) #otherwise does not read plot correctly
FG_sumAbund$sum <- as.numeric(FG_sumAbund$sum)
FG_sumAbund$plot <- as.factor(FG_sumAbund$plot)

# pivot-wider
SbyE_fg <- pivot_wider(FG_sumAbund, id_cols = "plot", 
                       names_from = c("troph"), values_from = "sum")
SbyE_fg <- as.data.frame(SbyE_fg)
rownames(SbyE_fg) <- SbyE_fg$plot

# creaete new and merge phyloseq
ps_new <- sample_data(SbyE_fg)
ps_FG <- merge_phyloseq(ps_land, ps_new)

# create proportional variables, relative to seq depth
sample_data(ps_FG)$prop_Patho_other <- c(sample_data(ps_FG)$Patho_other/sample_data(ps_FG)$seqDepth_onlyFungi)
sample_data(ps_FG)$prop_Symbio_other <- c(sample_data(ps_FG)$Symbio_other/sample_data(ps_FG)$seqDepth_onlyFungi)
sample_data(ps_FG)$prop_Other <- c(sample_data(ps_FG)$Other/sample_data(ps_FG)$seqDepth_onlyFungi)
sample_data(ps_FG)$prop_na <- c(sample_data(ps_FG)$na/sample_data(ps_FG)$seqDepth_onlyFungi)

# #visualize
# plot(log2(ps_FG@sam_data$prop_Patho_other) ~ log2(ps_FG@sam_data$prop_Symbio_other))
# plot(log2(ps_FG@sam_data$prop_Patho_other) ~ log2(ps_FG@sam_data$prop_Other))
# plot(log2(ps_FG@sam_data$prop_Symbio_other) ~ log2(ps_FG@sam_data$prop_Other))
# plot(log2(ps_FG@sam_data$prop_Patho_other) ~ log2(ps_FG@sam_data$prop_na))
# plot(log2(ps_FG@sam_data$prop_Symbio_other) ~ log2(ps_FG@sam_data$prop_na))
# plot(log2(ps_FG@sam_data$prop_Other) ~ log2(ps_FG@sam_data$prop_na))

#organize for linear models
SbyE_new <- data.frame(sample_data(ps_FG)@.Data)
colnames(SbyE_new) <- ps_FG@sam_data@names
rownames(SbyE_new) <- ps_FG@sam_data@row.names

```

# Models

### RRPP/MANOVA - Species-wide comparison
* Does not include Variety (would need to split by crop)

```{r, fig.width = 7}
par(mfrow = c(1,2))
hist(sqrt(SbyE_new$prop_Patho_other))
hist(log10(SbyE_new$prop_Symbio_other))
# should be transformed if run in a typical linear model. Likely fine for RRPP/permutation procedure

```

```{r, results = 'hide'}
fg_manova <- lm.rrpp(cbind(prop_Patho_other, prop_Symbio_other) ~ crop +
                          crop:site +
                          crop:site:plot.rep, iter = 999,
                          print.progress = FALSE, data = SbyE_new, SS.type="III")
```

```{r}
full.anova <- anova(fg_manova, effect.type = "F", error = c("crop:site:plot.rep",
    "crop:site:plot.rep", "Residuals"))
summary(full.anova, formula = false)
```

```{r, results = 'hide', echo = FALSE, fig.keep = 7}
path_bySite <- ggplot(SbyE_new, aes(y = prop_Patho_other, x = nestSite, fill = factor(nestSite))) +
    geom_boxplot(outlier.shape = NULL) + scale_fill_manual(values = site_color) +
    geom_jitter(width = .1)+
    scale_x_discrete("Site") +
    scale_y_continuous("Proportion of Pathogen Sequences", limits = c(-0.01,0.65)) + 
    guides(fill = 'none') +
    theme(axis.text.x = element_text(angle = 90))
symb_bySite <- ggplot(SbyE_new, aes(y = prop_Symbio_other, x = nestSite, fill = factor(nestSite))) +
    geom_boxplot(outlier.shape = NULL) + scale_fill_manual(values = site_color) +
    geom_jitter(width = .1)+
    scale_x_discrete("Site") +
    scale_y_continuous("Proportion of Symbiont Sequences", limits = c(-0.01,0.65)) + 
    guides(fill = 'none') +
    theme(axis.text.x = element_text(angle = 90))
multiplot(path_bySite, symb_bySite, cols = 2)
```

### CVs of Proportions across Sites
```{r}
patho_CV <- c(round(tapply(SbyE_new$prop_Patho_other, SbyE_new$nestSite,
    stats::sd)/tapply(SbyE_new$prop_Patho_other, SbyE_new$nestSite, mean),4)*100)

symbio_CV <- c(round(tapply(SbyE_new$prop_Symbio_other, SbyE_new$nestSite, 
    stats::sd)/tapply(SbyE_new$prop_Symbio_other, SbyE_new$nestSite, mean),4)*100)

data.frame(patho_CV, symbio_CV)
```

* These C.V.'s are quite high, above 100% in some cases. Indicates data really should be analyzed at plot level.
* *Dilemma* all of our follow-up analyses are at site level, such as climate and landscape vars. We'd risk pseudo-replication to analyze at plot level for data that only really exists at site level.

### example of pseudo-replication
```{r}

test <- lm(prop_Patho_other ~ LT_90day, data = SbyE_new)
Anova(test, type = "III")
summary(test)$coefficients
ggplot(SbyE_new, aes(x=LT_90day, y = prop_Patho_other, colour = nestSite)) +
    geom_point(size = 3) + scale_colour_manual(values = site_color)


```





#### end
```{r}
sessionInfo()
```

