---
title: "Landscape metrics"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
author: "Briana K. Whitaker"
date: "`r Sys.Date()`"
---
\fontsize{9}{10}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning=FALSE, message=FALSE)
```
---


* Run using `r version[['version.string']] `.

# Objective
This document reports on the landscape-level analyses from the NC Microbiome Landscapes Project, performed in the summer of 2019. 


```{r, include=FALSE} 
x<-c("ggplot2", "phyloseq", "dplyr", "RColorBrewer", "DESeq2", "viridis", "car", 'olsrr', 'reshape2')  
lapply(x, require, character.only = TRUE)

#add 'not in' function
`%nin%` = Negate(`%in%`)

#set seed
#runif(1, min = 0, max = 1000)
set.seed(213) 

# set ggplot2 theme
theme_set(theme_bw(base_size=16)) 
theme_update(panel.grid.major=element_line(0), panel.grid.minor=element_line(0))


source("./code/multiplot.function.R")

# color palettes

spp_color <- c("#E6AB02", "#D95F02", "#1B9E77", '#7570B3') 
#gold, orange, green, purple
#levels(sample_data(ps_filter)$species)
#"CRN"  "SOY"  "SWCH" "WHT" 
#https://mdigi.tools/color-shades/#a6761d
crn_color <- c('#fed561', '#fdc422', '#dda402', '#9e7501')
soy_color <- c('#fea561', '#fd8122', '#dd6102', '#9e4501')
swc_color <- c('#77e8c6', '#40deaf', '#21bf90', '#178867')
wht_color <- c('#cfcde4', '#8f8ac1', '#555096', '#2f2c53')

site_color <- c(crn_color, soy_color, swc_color, wht_color)

```


```{r, include=FALSE} 
## Phyloseq obj
load("./data/MBLand_ps_land.RData")
ps_land #595 ASVs
load("./data/MBLand_ps_onlyFungi_lulu_noCont.RData")
ps_lulu #1211 ASVs  -----> pre-singleton and low abundance taxa


# from non-filtered dataset, Remove reads with abundance <=2, but keep singletons (<2 based on analysis of controls)
ps_filt1 <- prune_taxa(taxa_sums(ps_lulu) > 2, ps_lulu)
ntaxa(ps_filt1)

alpha1 <- phyloseq::estimate_richness(ps_land, split=TRUE, measures="Observed")
alpha <- phyloseq::estimate_richness(ps_filt1, split=TRUE, measures="Observed")

row.names(alpha1)<-sapply(stringr::str_remove_all(rownames(alpha1),"X"),"[")
row.names(alpha)<-sapply(stringr::str_remove_all(rownames(alpha),"X"),"[")

sam <- phyloseq::sample_data(ps_land)
sam_rich1 <- merge(sam, alpha1, by="row.names", all=TRUE)
rownames(sam_rich1) <- rownames(sam)
sam_rich <- merge(sam_rich1, alpha, by.x="Row.names", by.y = "row.names", all=TRUE)
colnames(sam_rich) #Observed.x = richness no singles, Observed.y = richness with singles
#rename columns
colnames(sam_rich) -> my_cols
colnames(sam_rich) <- c(my_cols[-c(61:62)], "Rich_nosing" , "Rich_sing" )
colnames(sam_rich)

#write.csv(sam_rich, "./data/MBLand_rich_plots.csv")

```


```{r, include=FALSE}
gee <- read.csv("./data/MBLand_SiteData_GEE.csv", stringsAsFactors = TRUE)
#colnames(gee)
gee2 <- gee %>% dplyr::select("Species_Site", "Longitude", "Latitude", 
            "Perc_crop_10km", "Perc_veg_10km", "Perc_crop_1km", "Perc_veg_1km")

vst.means2 <- read.csv("./data/MBLand_CommDistance_VST_means.csv", 
                       row.names = 1, stringsAsFactors = TRUE)

# merge datasets, as in 4-MBLand_CommAnalyses code
vst_gee <- merge(vst.means2, gee2, by.x = 'site', by.y = 'Species_Site', all.x = TRUE, all.y = FALSE)
dim(vst_gee)

# subset sam_rich, remove tassel
sam_rich2 <- sam_rich %>% filter(leaf.type != "Tassel")
sam_rich2$nestSite <- as.factor(paste(sam_rich2$species, sam_rich2$site, sep ="-"))

#create a means, sd, se, and ci dataset
rich.means <- data.frame(nestSite = c(levels(sam_rich2$nestSite)),
  Rich_count = c(table(sam_rich2$nestSite)),
  Rich_nosing = c(tapply(sam_rich2$Rich_nosing, sam_rich2$nestSite, mean)), 
  Rich_nosing_sd = c(tapply(sam_rich2$Rich_nosing, sam_rich2$nestSite, sd)), 
  Rich_sing = c(tapply(sam_rich2$Rich_sing, sam_rich2$nestSite, mean)), 
  Rich_sing_sd = c(tapply(sam_rich2$Rich_sing, sam_rich2$nestSite, sd))  )

# add se, ci-95%, and 95% confidence interval
rich.means$Rich_nosing_se <- (rich.means$Rich_nosing_sd/sqrt(rich.means$Rich_count))
rich.means$Rich_sing_se <- (rich.means$Rich_sing_sd/sqrt(rich.means$Rich_count))

rich.means$Rich_ciMult <- qt(.975, rich.means$Rich_count-1)

rich.means$Rich_nosing_ci <- rich.means$Rich_nosing_se * rich.means$Rich_ciMult 
rich.means$Rich_sing_ci <- rich.means$Rich_sing_se * rich.means$Rich_ciMult 

#make factor
rich.means$nestSite <- factor(rich.means$nestSite)



# merge vst_gee with means data for sam_rich, then write out

rich <- merge(vst_gee, rich.means, by.x = 'site', by.y = 'nestSite', all.x = TRUE, all.y = TRUE)
dim(rich)
#write.csv(rich, "./data/MBLand_Rich_GEE_means.csv")
```

# Open File and Check Correlations
```{r}
# Open file of average richness ("Rich") by land cover data at 1 and 10km
RichLU <- read.csv("./data/MBLand_Rich_GEE_means.csv", stringsAsFactors=TRUE, header=TRUE, row.names = 1)
# count, mean, sd, se, ciMult, ci all refer to euclidean distance of the VST averaged per site
RichLU$species <- as.factor(unname(sapply(
  RichLU$site,  function(x) strsplit(as.character(x), "-")[[1]][1] )) )


env_corr <- subset(RichLU, select=c( Perc_crop_10km, Perc_veg_10km, Perc_crop_1km, Perc_veg_1km)) 
cor(env_corr)
# all are strongly correlated, likely to be collinear and will need to drop some from model

#tiff("./figures/Corrplot Landcover factors.tiff", width=7, height=4.36, units="in", res=600)
corrplot::corrplot(cor(env_corr), method="number", type="lower", order="hclust", #tl.pos = 'l',
                   tl.col = 'black', tl.cex = .9, cl.cex = 0.8, number.cex = .9)
#dev.off()
```

# Check Crop Differences in landcover
```{r}
p1 <- ggplot(RichLU, aes(y=Perc_crop_10km, x=species, color = species)) +
      geom_violin() + geom_jitter(size = 3, width = 0.05) + guides(color = "none") +
      scale_color_manual(values = spp_color) + 
      annotate(geom = "text", x = .7, y = 58, label = "C)", size = 5.5)
p2 <- ggplot(RichLU, aes(y=Perc_crop_1km, x=species, color = species)) +
      geom_violin() + geom_jitter(size = 3, width = 0.05) + guides(color = "none") +
      scale_color_manual(values = spp_color) + 
      annotate(geom = "text", x = .7, y = 95, label = "D)", size = 5.5)

p3 <- ggplot(RichLU, aes(y=Perc_veg_10km, x=species, color = species)) +
      geom_violin() + geom_jitter(size = 3, width = 0.05) + guides(color = "none") +
      scale_color_manual(values = spp_color) + 
      annotate(geom = "text", x = .7, y = 70, label = "A)", size = 5.5)
p4 <- ggplot(RichLU, aes(y=Perc_veg_1km, x=species, color = species)) +
      geom_violin() + geom_jitter(size = 3, width = 0.05) + guides(color = "none") +
      scale_color_manual(values = spp_color) + 
      annotate(geom = "text", x = .7, y = 51.3, label = "B)", size = 5.5)

#tiff("./figures/Distribution Landcover factors.tiff", width=6.5, height=5.5, units="in", res=600)
multiplot(p3,p4,p1,p2, cols = 2)
#dev.off()
```

# RICHNESS
## NO SINGLETONS
### Define regression model and check assumptions
```{r}
# Define initial model
model1 <- lm(Rich_nosing ~ Perc_crop_10km + Perc_veg_10km + Perc_crop_1km + Perc_veg_1km, data=RichLU)
Anova(model1)
olsrr::ols_plot_resid_fit(model1)

olsrr::ols_test_normality(model1) 

olsrr::ols_test_breusch_pagan(model1, rhs=TRUE, multiple=TRUE) 

olsrr::ols_coll_diag(model1)
```

```{r}
# vif's are 3-7, need to drop variables
# Redefine model with worst variable dropped
model2 <-lm(Rich_nosing ~ Perc_crop_10km + Perc_crop_1km + Perc_veg_1km, data=RichLU)
olsrr::ols_coll_diag(model2)
```

```{r}
# vif's are still 3-5
# drop next worst variable
model3 <- lm(Rich_nosing ~ Perc_crop_10km + Perc_veg_1km, data=RichLU)
olsrr::ols_coll_diag(model3)

# both factors have vif's under 2
```

### Regression no singletons
```{r}
olsrr::ols_step_best_subset(model3)

# best model based on AIC and SBC (Schwarz Bayesian Information Criterior or just BIC as far as I can tell) has only Perc_veg_1km; also has best Adj.R2
```


```{r}
# run simple linear regression on best model to get P value
olsrr::ols_regress(Rich_nosing ~ Perc_veg_1km, data=RichLU)

# contrast to P value for model3
olsrr::ols_regress(model3)
```


# RICHNESS
## WITH SINGLETONS
### Define regression model and check assumptions
```{r}
model4 <- lm(Rich_sing ~ Perc_crop_10km + Perc_veg_10km + Perc_crop_1km + Perc_veg_1km, data=RichLU)

olsrr::ols_plot_resid_fit(model4)

olsrr::ols_test_normality(model4) 

olsrr::ols_test_breusch_pagan(model4, rhs=TRUE, multiple=TRUE) 

olsrr::ols_coll_diag(model4)
```

```{r}
# vif's are 3-7, need to drop variables
# Redefine model with worst variable dropped
model5 <-lm(Rich_sing ~ Perc_crop_10km + Perc_crop_1km + Perc_veg_1km, data=RichLU)
olsrr::ols_coll_diag(model5)
```

```{r}
# vif's are still 3-5
# drop next worst variable
model6 <- lm(Rich_sing ~ Perc_crop_10km + Perc_veg_1km, data=RichLU)
olsrr::ols_coll_diag(model6)

# both factors have vif's under 2!
```

### Regression w/ singletons
```{r}
olsrr::ols_step_best_subset(model6)

# best model based on AIC, C(p) has only Perc_veg_1km; also has best Adj.R2
```


```{r}
# run simple linear regression on best model to get P value
olsrr::ols_regress(Rich_sing ~ Perc_veg_1km, data=RichLU)

# contrast to P value for model3
olsrr::ols_regress(model6)
```



# WITHIN SITE HETEROGENEITY
### Define regression model and check assumptions
```{r}
model7 <- lm(mean ~ Perc_crop_10km + Perc_veg_10km + Perc_crop_1km + Perc_veg_1km, data=RichLU)

olsrr::ols_plot_resid_fit(model7)

olsrr::ols_test_normality(model7) 

olsrr::ols_test_breusch_pagan(model7, rhs=TRUE, multiple=TRUE) 

olsrr::ols_coll_diag(model4)
```

```{r}
# vif's are 3-7, need to drop variables
# Redefine model with worst variable dropped
model8 <-lm(mean ~ Perc_crop_10km + Perc_crop_1km + Perc_veg_1km, data=RichLU)
olsrr::ols_coll_diag(model8)
```

```{r}
# vif's are still 3-5
# drop next worst variable
model9 <- lm(mean ~ Perc_crop_10km + Perc_veg_1km, data=RichLU)
olsrr::ols_coll_diag(model9)

# both factors have vif's under 2!
```

### Regression
```{r}
olsrr::ols_step_best_subset(model9)

# best model based on AIC, C(p) has only Perc_veg_1km; also has best Adj.R2
```


```{r}
# run simple linear regression on best model to get P value
olsrr::ols_regress(mean ~ Perc_veg_1km, data=RichLU)

# contrast to P value for model3
olsrr::ols_regress(model9)
```




# FIGURES
```{r}
# make long format data

names(RichLU)
head(RichLU)

RichLU_long <- melt(RichLU, id.vars = c('site', 'mean', 'ci', 
                    'Longitude', 'Latitude', 'Rich_nosing', 'Rich_nosing_ci',
                    'Rich_sing', 'Rich_sing_ci'), 
  measure.vars = c('Perc_crop_10km', 'Perc_veg_10km', 'Perc_crop_1km', 'Perc_veg_1km'),
  variable.name = 'landcover.type', value.name = 'percentage')

RichLU_long$scale <- as.factor(unname(sapply(RichLU_long$landcover.type,  function(x) strsplit(as.character(x), "_")[[1]][3] )) )
RichLU_long$type <- as.factor(unname(sapply(RichLU_long$landcover.type,  function(x) strsplit(as.character(x), "_")[[1]][2] )) )

levels(RichLU_long$type) <- c("Cropland", "Vegetation")

# then reverse order for figure
RichLU_long$type <- factor(RichLU_long$type, levels = c("Vegetation", "Cropland"))


```

```{r}
land_A <- ggplot(RichLU_long[RichLU_long$landcover.type == "Perc_veg_1km",],
       aes(x = percentage, y = Rich_nosing, colour = site)) +
  geom_smooth(aes(x = percentage, y = Rich_nosing), method=lm, formula = y ~ x,
              se = TRUE, color = '#666666', linetype = 1) +
  geom_point(size = 4) +
  geom_errorbar(aes(ymin=Rich_nosing-Rich_nosing_ci, ymax=Rich_nosing+Rich_nosing_ci), width=.8) +
  scale_y_continuous("Avg. Within Site Richness", limits = c(0,90)) +
  scale_x_continuous("% Unmanaged Vegetation\n1-km scale", limits = c(0,55)) +
  scale_color_manual("", values = site_color) +
  theme(legend.position="none", 
        axis.title = element_text(size = 28), axis.text = element_text(size = 28)) +
  annotate(geom = "text", x = 3, y = 84.65, label = "A)", size = 9)

land_B <- ggplot(RichLU_long[RichLU_long$landcover.type == "Perc_veg_1km",],
       aes(x = percentage, y = mean, colour = site)) +
  geom_smooth(aes(x = percentage, y = mean), method=lm, formula = y ~ x,
              se = TRUE, color = '#666666', linetype = 1) +
  geom_point(size = 4) +
  geom_errorbar(aes(ymin=mean-ci, ymax=mean+ci), width=.8) +
  scale_y_continuous("Avg. Within Site Dissimilarity", limits = c(0,65)) +
  scale_x_continuous("% Unmanaged Vegetation\n1-km scale", limits = c(0,55)) +
  scale_color_manual("", values = site_color) +
  theme(legend.position="none", 
        axis.title = element_text(size = 28), axis.text = element_text(size =28)) +
  annotate(geom = "text", x = 3, y = 61, label = "B)", size = 9)


#tiff("./figures/Dissim & Richness (Alt) ~ Perc_Veg.tiff", width=13.2, height=7.04, units="in", res=600)
multiplot(land_A, land_B, cols=2)
#dev.off()

```



```{r}
sessionInfo() 
```












